Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering

### 2016-07-12 paper/structureScene.tex
address joakim's comments in IV. B

\subsection{Invariance and stability in acoustic scene classification}
The notion of invariance to time shifts plays an essential role in auditory scene classification.
Indeed, recordings may be shifted locally without affecting their perception and therefore such shifts do not convey any information about the class.
To discard this superfluous source of variability, signals can be mapped to a time-shift invariant feature space before training the classifier, eliminating the need for the classifier to explicitly learn this invariance.
From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ \ja{$x_1$, $\gamma_1$, etc have not been introduced yet.} of a signal $\boldsymbol{x}(t)$ --
a representation invariant to time shifts shorter in duration than $T$ can be obtained
by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
set to $1/T$ \ja{How are we defining frequency here, as $\omega$ or $2\pi\omega$? In the former case, we should have $2\pi/T$. Probably we should define our Fourier transform somewhere to nail this down}:
\begin{equation}
\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = (\boldsymbol{x_1} \ast \boldsymbol{\phi}) (t).
\end{equation}
A downside is that the transient information in $\boldsymbol{x_1}$ at scales finer than $T$ are lost by this low-pass filtering, reducing discriminability in feature space.
To address this issue, the temporal scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$ and then applying complex modulus.

By resorting to wavelet transform modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to time-warping deformation,
in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}.
In addition time-shift invariance, this stability property is crucial to signal classification, since it guarantees robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.

@@

\subsection{Temporal scattering}
Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$. \ja{I mentioned this already, but it would be good to define the Fourier transform so that we can make these statements more specific}
A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
according to a geometric sequence of scales $2^{\gamma_1/Q_1}$, obtaining
\begin{equation}
\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(2^{-\gamma_1/Q_1} t)\mbox{.}
\end{equation}
The variable $\gamma_1$ is an inverted log-frequency, taking integer values between $0$ and $(J_1 \times Q_1 - 1)$.
In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$ (close to the Nyquist frequency of the audio recordings) the number of octaves $J_1$ to $10$ (the lower end of human hearing range) and the number of wavelets per octave $Q_1$ to $8$.
For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$, a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$, and a quality factor of $Q_1$.

>>>>

\subsection{Invariance and stability in acoustic scene classification}
The notion of invariance to time shifts plays an essential role in auditory scene classification.
Indeed, recordings may be shifted locally without affecting their perception and therefore such shifts do not convey any information about the class.
To discard this superfluous source of variability, signals can be mapped to a time-shift invariant feature space before training the classifier, eliminating the need for the classifier to explicitly learn this invariance.
From any set of descriptors $\boldsymbol{x_1}(t,\gamma)$, where $\gamma$ denotes a descriptor index, a representation invariant to time shifts shorter in duration than $T$ can be obtained by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency set to $1/T$, as measured in Hertz:
\begin{equation}
\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = (\boldsymbol{x_1} \ast \boldsymbol{\phi}) (t).
\end{equation}
A downside is that the transient information in $\boldsymbol{x_1}$ at scales finer than $T$ are lost by this low-pass filtering, reducing discriminability in feature space.
To address this issue, the temporal scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$ and then applying complex modulus.

By resorting to wavelet transform modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to time-warping deformation,
in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}.
In addition to invariance, this stability property is crucial to signal classification, since it guarantees robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.

@@

\subsection{Temporal scattering}
Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int x(t) \exp(\mathrm{i} 2\pi t \omega) \; \mathrm{d}\omega$. Note that the frequency variable $\omega$ is expressed in Hertz, not in radians per second.

Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$.
A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
according to a geometric sequence of scales $2^{\gamma_1/Q_1}$, obtaining
\begin{equation}
\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(2^{-\gamma_1/Q_1} t)\mbox{.}
\end{equation}
The variable $\gamma_1$ is the opposite number of a log-frequency, taking integer values between $0$ and $(J_1 \times Q_1 - 1)$.
In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$ (close to the Nyquist frequency of the audio recordings) the number of octaves $J_1$ to $10$ (the lower end of human hearing range) and the number of wavelets per octave $Q_1$ to $8$.
For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$, a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$, and a quality factor of $Q_1$.
